---
title: "Lab4"
output: html_document
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
#remove everything from workspace
rm(list = ls(all = TRUE))
# tell R to use the packages you installed
library(MendelianRandomization)
library(metafor)
```


# 1. read in the leadihd SNPs file you saved from lab 1 
```{R}
read_csv("Lab1/leadihd21.csv") -> pl
nrow(pl) # 13
```


# 2 get estimates for ihd using Mendelian randomization package with fixed effects,random effects, weighted median and MR-Egger

```{R}
mr_input(pl$beta,
         pl$se,
         pl$beta.outcome,
         pl$se.outcome) -> MRInputObject

list(IVW_fixed = mr_ivw(MRInputObject,
                        model = "fixed"),
     IVW = mr_ivw(MRInputObject), # IVW with multiplicative random effects
     `Weighted median method` = mr_median(MRInputObject),
     `MR-Egger` = mr_egger(MRInputObject)) -> MR_list

MR_list[[4]] -> df
df

```

## How do you interpret the following information for MR-Egger:
* IVW, -0.017
* Weighted median
  * 0.001
* MR-Egger estimate?
  * 0.051
* The MR-Egger intercept?
  * -0.011 != 0
  * If the MR-Egger intercept is nonzero, then the IVW estimate is invalid
* The Cochran Q: MR-Egger uses Cochran Q to indicate over dispersion (i.g., heterogeneity)
  * 44.2113 (p-value = 0.0000) -> High heterogeneity
* IGX2: MR-Egger uses indicate instrument strength
  * 22.0%


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{R}
rm(list=ls(all = TRUE))

#install MRPRESSO
devtools::install_github("rondolab/MR-PRESSO")
library(MRPRESSO)
library(MendelianRandomization)
library(metafor)
# read in file of SNPs to exposure and outcome for lead
pl<-read.csv("Lab1/leadihd21.csv",header=TRUE)
nrow(pl)
```


```{R}
# run MR-Presso 
mr_presso(BetaOutcome = "beta.outcome",
          BetaExposure = "beta",
          SdOutcome = "se.outcome",
          SdExposure = "se",
          OUTLIERtest = TRUE,
          DISTORTIONtest = TRUE,
          data = pl,
          NbDistribution = 10000, 
          SignifThreshold = 0.05)
```

$`Main MR results`

$`MR-PRESSO results`
$`MR-PRESSO results`$``
$`MR-PRESSO results`$`Global Test`$
[1] 

* Global Test
  * Residual sum and outlier, ideally <0.1
  * RSSobs = 51.7629
* Pvalue <1e-04
* Outlier, y
* Distortion Coefficient, whether the new estimate is
different from the original estimate, beta = -195.7128 (p = 0.2006)


# identify the SNPs MR-Pressos said were outliers
```{R}

```


# lets have a look at the original forest plot again
# 4 get forest plot with fixed effects, same as Mendelianrandomization IVW with fixed effects 

```{R}
x <- pl$beta # beta for SNP to exposure
sigmax <- pl$se # its standard errors
y <- pl$beta.outcome # beta for SNP to outcome
sigmay <- pl$se.outcome # its standard errors
pl$Wald <- y/x #Wald estimate
pl$Waldvar <- (sigmay^2/x^2) # using Burgess's method
pl$lab <- paste(pl$SNP, pl$gene, sep=" ")

pl <- pl[pl$SNP != pl[7,]$SNP,
         ] #uncomment this row to exclude rs550057

dmres <- rma.uni(yi = pl$Wald,
                 vi = pl$Waldvar,
                 slab = pl$lab,
                 method = "FE")
dmres
forest(dmres,
       atransf = exp,
       xlab=" ",
       mlab = "Ischemic heart disease (OR)",
       at = log(c(.5, 1,2)),
       xlim=c(-1.7,1.3),
       cex=.8)
```

#2 get estimates for ihd using Mendelian randomization package with fixed effects,random effects, weighted median and MR-Egger
```{R}
mr_input(pl$beta,
         pl$se,
         pl$beta.outcome,
         pl$se.outcome) -> MRInputObject

mr_allmethods(MRInputObject,
              method = "all")
mr_conmix(MRInputObject)
mr_lasso(MRInputObject)
```
